Learning and comparing functional connectomes between populations

### 2012-12-24 paper.tex
ENH: paragraph on covariance/correlation matrix

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation -- 
Upper left: regressing out CompCor signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the default mode network.
% XXX: should I outline the default mode network on the main matrix?
-- Upper right: regressing out only movement parameters. -- Lower right:
regressing out movement parameters and global signal mean.
% XXX: should this remark go here?
No frequency filtering was applied here.
\label{fig:correlation_matrices}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

\paragraph{Correlation and partial correlations}
%
Functional connectivity between the ROIs can be measured by computing the
correlation matrix of the extracted signals. An important and often
neglected point is that the sample correlation matrix, \emph{i.e.} the
correlation matrix obtained by plugging the observed signal in the
correlation matrix formula, is not the population correlation matrix,
\emph{i.e.} the correlation matrix of the data-generating process. If the
number of measurements was infinite, the two would coincide, however if
this number is not large compared to the number of connection (that
scales as the square of the number of ROIs), the sample correlation
matrix is a poor estimate of the underlying population correlation matrix.
In other words, the sample correlation matrix captures a lot of sampling
noise, intrinsic randomness that arises in the estimation of correlations
from short time series. Conclusions drawn from the sample correlation
matrix can easily reflect this estimation error.

Varoquaux \emph{et al.} \cite{varoquaux2010c} and Smith \emph{et al.}
\cite{smith2011} have shown respectively on rest fMRI and on realistic
simulations that a good choice of correlation matrix estimator could
recover the connectivity structure, where the sample correlation matrix
would fail. In general, the choice of a better estimate depends on the
settings and the end goals \cite{varoquaux2012,varoquaux2010b}, however
the Ledoit-Wolf shrinkage estimate \cite{ledoit2004} is a simple,
computationally-efficient, and parameter-less alternative that performs
uniformly better than the sample correlation matrix
\cite{varoquaux2012,varoquaux2010c} and should always be prefered.

Remark: difference between covariance and correlation matrix.

partial correlation, sparse iCov. \cite{smith2011,varoquaux2010b}.

Give intuition on partial correlation/inverse covariance.

Remark that partial correlations consistently have no negative values.

Variety
of estimation strategies for sparse iCov, that have an impact on the
resulting network \cite{varoquaux2012}\footnote{Better estimation
algorithm is recent Honorio paper.}. Discuss setting the regularization
parameter.

@@

    \includegraphics[width=\linewidth]{group_emp_prec_global_mean.pdf}%
\end{minipage}%

\caption{
Inverse-covariance matrices
}
\end{figure}
%---------------------------------------------------------------------------}}}

% XXX: should we have a summary paragraph commenting on the structure of
% the graphs and correlation matrices extracted.
% And also commenting on the meaning of regressions and sign of
% correlation coefficients (the whole global mean debate).
\cite{chai2011} (note: chai2011 used compcor)

>>>>

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation -- 
\textbf{Upper left}: regressing out CompCor signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the default mode network.
% XXX: should I outline the default mode network on the main matrix?
-- \textbf{Upper right}: regressing out only movement parameters. -- 
\textbf{Lower right}:
regressing out movement parameters and global signal mean.
% XXX: should this remark go here?
No frequency filtering was applied here.
\label{fig:correlation_matrices}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

\paragraph{Correlation and partial correlations}
%
Functional connectivity between the ROIs can be measured by computing the
correlation matrix of the extracted signals. An important and often
neglected point is that the sample correlation matrix, \emph{i.e.} the
correlation matrix obtained by plugging the observed signal in the
correlation matrix formula, is not the population correlation matrix,
\emph{i.e.} the correlation matrix of the data-generating process. If the
number of measurements was infinite, the two would coincide, however if
this number is not large compared to the number of connection (that
scales as the square of the number of ROIs), the sample correlation
matrix is a poor estimate of the underlying population correlation
matrix. In other words, the sample correlation matrix captures a lot of
sampling noise, intrinsic randomness that arises in the estimation of
correlations from short time series. Conclusions drawn from the sample
correlation matrix can easily reflect this estimation error.
%
Varoquaux \emph{et al.}~\cite{varoquaux2010c} and Smith \emph{et
al.}~\cite{smith2011} have shown respectively on rest fMRI and on
realistic simulations that a good choice of correlation matrix estimator
could recover the connectivity structure, where the sample correlation
matrix would fail. In general, the choice of a better estimate depends on
the settings and the end goals \cite{varoquaux2012,varoquaux2010b},
however the Ledoit-Wolf shrinkage estimate \cite{ledoit2004} is a simple,
computationally-efficient, and parameter-less alternative that performs
uniformly better than the sample correlation matrix
\cite{varoquaux2012,varoquaux2010c} and should always be preferred.

@@

    \includegraphics[width=\linewidth]{group_l21_prec_global_mean.pdf}%
\end{minipage}%

\caption{
Inverse-covariance matrices:
\label{fig:icov}
}
\end{figure}
%---------------------------------------------------------------------------}}}

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables -- strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on normalized signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
graphical-lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is a good general-purpose approach. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see Fig\,\ref{fig:icov}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to set this parameter is via cross-validation
\cite{varoquaux2010c}.

% XXX: should we have a summary paragraph commenting on the structure of
% the graphs and correlation matrices extracted.
% And also commenting on the meaning of regressions and sign of
% correlation coefficients (the whole global mean debate).
% \cite{chai2011} (note: chai2011 used compcor)
% Remark that partial correlations consistently have no negative values.
